### 2024-05-06 lattice_based_attacks.tex
version 2, fix grammar

\documentclass[a4paper,12pt]{report}

\usepackage{alltt, fancyvrb, url}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{wrapfig}
% \usepackage{algorithm}% http://ctan.org/pkg/algorithms
% \usepackage[ruled, vlined]{algorithm2e}
% \usepackage{algorithm}
\usepackage[]{algorithm2e}
\usepackage{titling}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{amsmath, stmaryrd, mathtools, algorithm}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{url}

\usepackage[english]{cleveref}

\usepackage[english]{babel}

\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{property}{Property}[section]
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\dd}{\cdot}
\newcommand{\subtitle}[1]{
  \posttitle{
    \par\end{center}
    \begin{center}\large#1\end{center}
    \vskip0.5em}
}

\title{Introduction to Lattice-based Attacks}
\subtitle{Version 2}

\author{Giovanni Di Santi}
\date{}

\begin{document}

\maketitle

\tableofcontents

\chapter{Introduction}

Lattices are potent mathematical structures widely utilized in computer science and mathematics to solve a broad spectrum of problems.
In recent years, numerous post-quantum cryptographic protocol candidates have employed the challenging problems associated with hard lattices,
for example, the \textbf{Learning With Errors} \cite{LWE}.
Additionally, lattice constructions are leveraged in cryptanalysis to compromise cryptographic schemes, typically through the use of
reduction algorithms such as \textbf{LLL} \cite{LLL}.

To comprehend the mathematics of lattices, a foundational understanding of linear algebra is essential. Therefore, the second chapter provides a summary of the
basic properties necessary for this understanding. We primarily referenced ``An Introduction to Mathematical Cryptography''\cite{mathcrypto14} to
elucidate these concepts, while also providing some geometric intuition behind certain operations.

Cryptographic protocols such as \textbf{RSA} and \textbf{ECDSA} are considered secure; however, various attacks can exploit a
relaxed model of each, e.g., what if some bits of the secret key are known? What if a portion of the message is known prior to encryption?
We introduce lattice-based attacks on these protocols and provide code in our repository \cite{repo} to verify the effectiveness of these attacks. The main reference for this discussion is
``Recovering cryptographic keys from partial information, by example''\cite{cryptoeprint:2020:1506}.

% Lattices are powerful mathematical objects that can be used in computer science and mathematics to solves an extensive range of different problems.
% In the recent years many post-quantum cryptographic protocols candidates involve the use of hard lattices problems,
% for example the \textbf{Learning With Errors} \cite{LWE}.
% Lattice constructions can also be used in cryptanalysis to break cryptographic schemes, typically using reduction
% algorithms as \textbf{LLL} \cite{LLL}.
% 
% To introduce the mathematics of lattices a basic linear algebra background is needed, so the second chapter is a summary of the
% basics properties needed to understand it. We used as main resource ``An Introduction to Mathematical Cryptography''\cite{mathcrypto14} to
% explain them, but we also tried to give some geometrical intuition behind some operations.
% 
% Cryptographic protocol as \textbf{RSA} and \textbf{ECDSA} are considered secure, however there are various attacks against a
% relaxed-model of each of them, i.e., What if we know some bits of the secret key? What if we know a part of the message before the encryption?
% We present an introduction to the attacks which involves lattices against these protocols. The code used to confirm the correctness
% of the attacks is available on the repository \cite{repo} and the main reference was
% ``Recovering cryptographic keys from partial information, by example''\cite{cryptoeprint:2020:1506}.

\chapter{Linear Algebra Background}

This chapter introduces key definitions and properties essential for understanding lattices. It progresses to cover the mathematics of lattices and related complex problems. By the end of this chapter, the reader should be well-equipped to follow the mathematical discussions in subsequent chapters.

\section{Vector Spaces}

\begin{definition}
    \textbf{Vector space}.
\end{definition}
A \textit{vector space} $V$ over $\R^{m}$ is a set closed under finite vector addition and scalar multiplication. This closure is characterized by:

\begin{center}
   $a_1v_1 + a_2v_2 \in V$ for any $v_1,v_2 \in V$ and $a_1,a_2 \in \R$
\end{center}

\begin{definition}
    \textbf{Linear Combinations}
\end{definition}

Consider any vectors $v_1, v_2, \ldots, v_k \in V$. A \textit{linear combination} of these vectors is any vector expressed as:

\begin{center}
    $\alpha_1v_1 + \alpha_2v_2 + \cdots + \alpha_kv_k$ where $\alpha_1, \ldots, \alpha_k \in \R$
\end{center}

\begin{definition}
    \textbf{Linear Independence}
\end{definition}

A set of vectors $v_1, v_2, \ldots, v_k \in V$ is \textit{linearly independent} if the only solution to the equation

\begin{center}
    $a_1v_1 + a_2v_2 + \cdots + a_kv_k = 0$
\end{center}

is $a_1 = a_2 = \cdots = a_k = 0$.

\begin{definition}
    \textbf{Bases}
\end{definition}

A set of linearly independent vectors $b = (v_1, \ldots, v_n) \in V$ constitutes a \textit{basis} of $V$ if every vector $w \in V$ can be represented as:

\begin{center}
    $w = a_1v_1 + a_2v_2 + \cdots + a_nv_n$
\end{center}

\begin{definition}
    \textbf{Vector's Length}
\end{definition}

The \textit{Euclidean norm} or length of a vector $v = (x_1, x_2, \ldots, x_m)$ is given by:

\begin{center}
    $\lVert v \rVert = \sqrt{x_1^2 + x_2^2 + \cdots + x_m^2}$
\end{center}

% \chapter{Linear Algebra Background}
% 
% We start with some definitions and properties needed to understand the lattices, to move later on the mathematics behind them and some
% hard problems related to it. At the end of the chapter the reader should be able to follow the math used in the other chapters.
% 
% \section{Vector Spaces}
% 
% \begin{definition}
%     \textbf{Vector space}.
% \end{definition}
% A \textit{vector space} $V$ is a subset of $\R^{m}$ which is closed under finite vector addition and scalar multiplication, with the property that
% 
% \begin{center}
%    $a_1v_1 + a_2v_2 \in V$ for all $v_1,v_2 \in V$ and all $a_1,a_2 \in \R$
% \end{center}
% 
% \begin{definition}
%     \textbf{Linear Combinations}
% \end{definition}
% 
% Let $v_1,v_2,\ldots,v_k \in V$. A \textit{linear combination} of $v_1,v_2,\ldots,v_k \in V$ is any vector of the form:
% 
% \begin{center}
%     $\alpha_1v_1 + \alpha_2v_2 + \cdots + \alpha_kv_k$ with $\alpha_1, \ldots, \alpha_k \in \R$
% \end{center}
% 
% \begin{definition}
%     \textbf{Linear Independece}
% \end{definition}
% 
% A set of vectors $v_1,v_2,\ldots,v_k \in V$ is \textit{linearly independent} if the the only way to get
% 
% \begin{center}
%     $a_1v_1 + a_2v_2 + \cdots + a_kv_k = 0$
% \end{center}
% 
% is to have $a_1 = a_2 = \cdots = a_k = 0$.
% 
% \begin{definition}
%     \textbf{Bases}
% \end{definition}
% 
% Taken a set of linearly independent vectors $b = (v_1,\ldots,v_n) \in V$ we say that $b$ is a \textit{basis} of $V$ if $\forall w \in V$ we can write
% 
% \begin{center}
%     $w = a_1v_1 + a_2v_2 + \cdots + a_nv_n$
% \end{center}
% 
% \begin{definition}
%     \textbf{Vector's length}
% \end{definition}
% 
% The vector's length or \textit{Euclidean norm} of $v = (x_1, x_2, \ldots, x_m)$ is
% 
% \begin{center}
%     $\lVert v \rVert = \sqrt{x_1^2 + x_2^2 + \cdots + x_m^2}$
% \end{center}

\begin{definition}
    \textbf{Dot Product}
\end{definition}

Given vectors $v, w \in V \subset \R^m$ where $v = (x_1, x_2, \ldots, x_m)$ and $w = (y_1, y_2, \ldots, y_m)$, the \textit{dot product} of $v$ and $w$ is defined as:

\begin{center}
    $v \cdot w = x_1y_1 + x_2y_2 + \cdots + x_my_m$

    or

    $v \cdot w = \lVert v \rVert \lVert w \rVert \cos{\theta}$
\end{center}

where $\theta$ is the angle between $v$ and $w$ when their starting points are positioned at the origin $O$.

\begin{figure}[!b]
    \centering
    \includegraphics[scale=0.2]{./img/dot_product.png}
    \caption{Dot Product By 3Blue1Brown \cite{3b1b}}
    \label{fig:dot_product}
\end{figure}

Geometrically, $v \cdot w$ represents the length of the projection of $w$ onto $v$ multiplied by the length of $v$, as illustrated in \ref{fig:dot_product}.

\begin{definition}
    \textbf{Orthogonal Basis}
\end{definition}

An \textit{orthogonal basis} for a vector space $V$ consists of vectors $v_1, \ldots, v_m$ such that:

\begin{center}
    $v_i \cdot v_j = 0$ for all $i \neq j$
\end{center}

If $\lVert v_i \rVert = 1$ for each $i$, then the basis is termed \textit{orthonormal}.

\begin{algorithm}[H]
    \vspace*{5px}
    $v_1^* = v_1$\;
    \\
    \For{$i \gets 2$ \KwTo $n$}{
        $\mu_{ij} = v_i \cdot v_j^* / \lVert v_j^* \rVert ^ 2$ for $1 \le j < i$ \;
        \\
        $v_i^* = v_i - \sum_{j=1}^{i-1} \mu_{ij}v_j^*$ \;
    }
    \caption{Gram-Schmidt Algorithm}
    \label{alg:gram_schmidt}
\end{algorithm}

Given a basis $b = (v_1, \ldots, v_n)$ of a vector space $V \subset \R^m$, the Gram-Schmidt algorithm generates an orthogonal basis $b^* = (v_1^*,\ldots,v_n^*)$. These two bases satisfy the property that Span$\{v_1,\ldots,v_i\}$ = Span$\{v_1^*,\ldots,v_i^*\}$ for all $i = 1,2,\ldots,n$.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.5\textwidth]{./img/gram_schmidt.png}
    \caption{Gram Schmidt Orthogonalization}
    \label{fig:gram_schmidt}
\end{figure}

Using vectors $v_1=(4, 1)$ and $v_2=(2, 3)$ as a basis and applying the Gram-Schmidt process results in $u_1=v_1=(4, 1), u_2=(-10/17, 40/17)$, as depicted in \ref{fig:gram_schmidt}.

\begin{definition}
    \textbf{Determinant}
\end{definition}

The \textit{determinant} of a square matrix is a scalar value that characterizes the matrix and is defined by the following properties:

\begin{enumerate}
    \item The determinant is linear in each row.
    \item The determinant changes sign when two rows are swapped.
    \item The determinant of the identity matrix is 1.
\end{enumerate}

A matrix is termed \textbf{singular} if its determinant is $0$, indicating it lacks an inverse.\\

Animations created by Grant Sanderson provide a fantastic visual understanding of the geometric intuition behind the \textit{determinant}\cite{determinant}.

% \begin{definition}
%     \textbf{Dot Product}
% \end{definition}
% 
% Let $v, w \in V \subset \R^m$ and $v = (x_1, x_2, \ldots, x_m), w = (y_1, y_2, \ldots, y_m)$, the \textit{dot product} of $v$ and $m$ is
% 
% \begin{center}
%     $v \dd m = x_1y_1 + x_2y_2 + \cdots + x_my_m$
% 
%     or
% 
%     $v \dd m = \lVert v \rVert \lVert w \rVert \cos{\theta}$
% \end{center}
% 
% where $\theta$ is the angle between $v$ and $w$ if we place the starting points of the vectors at the origin $O$.
% 
% \begin{figure}[!b]
%     \centering
%     \includegraphics[scale=0.2]{./img/dot_product.png}
%     \caption{Dot Product By 3Blue1Brown \cite{3b1b}}
%     \label{fig:dot_product}
% \end{figure}
% 
% Geometrically speaking $v \cdot m$ is the length of $w$ projected to $v$ multiplied by the length of $v$ as shown in \ref{fig:dot_product}
% 
% \begin{definition}
%     \textbf{Ortoghonal Basis} 
% \end{definition}
% 
% An \textit{ortoghonal basis} for a vector space $V$ is a basis $v_1, \ldots, v_m$ with the property that
% 
% \begin{center}
%     $v_i \dd v_j = 0$ for all $i \neq j$
% \end{center}
% 
% If $\lVert v_i \rVert = 1$ for all $i$ then the basis is \textit{orthonormal}.
% 
% \begin{algorithm}[H]
%     \vspace*{5px}
%     $v_1^* = v1$\;
%     \\
%     \For{$i \gets 2$ \KwTo $n$}{
%         $\mu_{ij} = v_i \cdot v_j^* / \lVert v_j^* \rVert ^ 2$ for $1 \le j < i$ \;
%         \\
%         $v_i^* = v_i - \sum_{j=1}^{i-1} \mu_{ij}v_j^*$ \;
%     }
%     \caption{Gram-Schmidt Algorithm}
%     \label{alg:gram_schmidt}
% \end{algorithm}
% 
% Let $b = (v_1, \ldots, v_n)$, be a basis for a vector space $V \subset \R^m$. There is an algorithm to create an orthogonal basis
% $b^* = (v_1^*,\ldots,v_n^*)$.
% The two bases have the property that Span$\{v_1,\ldots,v_i\}$ = Span$\{v_1^*,\ldots,v_i^*\}$ for all $i = 1,2,\ldots,n$
% 
% \begin{figure}[htpb]
%     \centering
%     \includegraphics[width=0.5\textwidth]{./img/gram_schmidt.png}
%     \caption{Gram Schmidt orthogonalization}
%     \label{fig:gram_schmidt}
% \end{figure}
% 
% If we take $v_1=(4, 1), v_2=(2, 3)$ as basis and apply gram schmidt we obtain $u_1=v_1=(4, 1), u_2=(-10/17, 40/17)$ as shown in \ref{fig:gram_schmidt}
% 
% \begin{definition}
%     \textbf{Determinant}
% \end{definition}
% 
% The \textit{determinant} of a square matrix is a function that satisfies the following properties:
% 
% \begin{enumerate}
%     \item The determinant is linear in each row.
%     \item The determinant reverses sign if two rows are interchanged.
%     \item The determinant of the identity matrix is equal to 1.
% \end{enumerate}
% 
% If the determinant of a matrix is $0$ then the matrix is called \textbf{singular} (without an inverse).\\
% 
% Grant Sanderson created fantastic animations to understand the geometrical intuition behind the \textit{determinant}\cite{determinant}.

% \section{Lattices}
% 
% \begin{definition}
%     \textbf{Lattice}
% \end{definition}
% 
% Let $v_1,\ldots,v_n \in \R^m, m \ge n$ be linearly independent vectors. A \textit{Lattice} $L$ spanned by $\{v_1,\ldots,n_n\}$ is the set of 
% all integer linear combinations of $v_1,\ldots,v_n$.
% 
% \begin{center}
%     $L = \bigg\{\sum_{i=1}^{n} a_iv_i, a_i \in \Z \bigg\}$
% \end{center}
% 
% If $v_i$ for every $i = 1,\ldots\,n$ has integer coordinates then the lattice is
% called \textit{Integral Lattice}.
% 
% On the figure \ref{fig:lattice0} we show a lattice $L$ with bases $v=(3, 1)$ and $w=(-1, 1)$, and on \ref{fig:lattice1} the same lattice $L$ with
% a different basis.
% 
% \begin{figure}[!bp]
%     \begin{minipage}[b]{0.50\textwidth}
%         \includegraphics[width=\textwidth]{./img/lattice_b0.png}
%         \caption{Lattice $L$ spanned by $v, w$}
%         \label{fig:lattice0}
%     \end{minipage}
%     \hspace{\fill}
%     \hspace{\fill}
%     \hspace{\fill}
%     \hspace{\fill}
%     \hspace{\fill}
%     \begin{minipage}[b]{0.50\textwidth}
%         \includegraphics[width=\textwidth]{./img/lattice_b1.png}
%         \caption{Lattice $L$ spanned by $v', w'$}
%         \label{fig:lattice1}
%     \end{minipage}
% \end{figure}
% 
% \clearpage
% 
% \section{Lattice Problems}
% 
% \subsection{SVP}
% 
% \textbf{The Shortest Vector Problem} (\texttt{SVP}): \textit{Find a nonzero vector $v \in L$ that minimez the Euclidean norm $\lVert v \rVert$.}
% 
% Gauss's developed an algorithm to find an optimal basis for a two-dimensional lattice given an arbitrary basis. The output of the algorithm
% gives the shortest nonzero vector in $L$ and in this way solves the \texttt{SVP}.
% 
% \begin{figure}[!b]
%     \centering
%     \includegraphics[width=0.6\textwidth]{./img/gauss_svp.png}
%     \caption{Gauss reduction}
%     \label{fig:gauss_svp}
% \end{figure}
% 
% \begin{algorithm}[H]
%     \vspace*{5px}
%     \While{true}{
%         \If{$\lVert v_2 \rVert < \lVert v_1 \rVert$}{
%             swap $v_1, v_2$ \;
%         }
%         m = $\left \lfloor{(v_1 \cdot v_2) {\lVert v_1 \rVert}^{-2}} \right \rfloor$ \\
%         \If{$m = 0$}{
%             return $v_1, v_2$ \;
%         }
%         $v_2 = v_2 - mv_1$ \\
%     }
%     \caption{Gauss Basis Reduction}
% \end{algorithm}
% 
% \textbf{Example}. Let $L$ a lattice spanned by $v_1 = (8, 4), v_2 = (7, 5)$, if we apply the gauss reduction algorithm we
% obtain $w_1 = (1, -1), w_2 = (6, 6)$ \ref{fig:gauss_svp}. $w_1$ is the shortest nonzero vector in the lattice $L$.
% 
% The bigger the dimension of the lattice, the harder is the problem and there isn't a polynomial algorithm to solve it.
% 
% \subsection{CVP}
% 
% \textbf{The Closest Vector Problem} (\texttt{CVP}): \textit{Given a vector $t \in \R^m$ that is not in $L$, find the vector $v \in L$
%     closest to $t$, in other words find a vector $v \in L$ that minimizes the Euclidean norm $\lVert t - v \rVert$.}
% 
% \begin{figure}[!ht]
%     \centering
%     \includegraphics[width=0.6\textwidth]{./img/cvp.png}
%     \caption{CVP}
%     \label{fig:cvp}
% \end{figure}
% 
% \textbf{Example}. Let $L$ a lattice spanned by $v_1 = (0, 1), v_2 = (2, 0)$, and a target vector $t = (2.3, 2.4)$. The closest vector
% to $t$ in the lattice $L$ is $w = (2, 2)$.\\
% 
% Both \textbf{CVP} and \textbf{SVP} are $\mathcal{NP}$-hard problems, thus the complexity of the problem increments with the dimension
% of the lattice. They can be used as primitive to create a trapdoor functions for a cryptographic scheme.
% 
% There are various algorithms that can output a solution under certain constraints for an approximation of both \textbf{SVP}
% and \textbf{CVP}, for example \textbf{LLL}\cite{LLL} and \textbf{Babai's algorithm}\cite{babai} respectively.

\section{Lattices}

\begin{definition}
    \textbf{Lattice}
\end{definition}

Consider $v_1, \ldots, v_n \in \R^m$ with $m \ge n$, which are linearly independent vectors. A \textit{lattice} $L$, spanned by $\{v_1, \ldots, v_n\}$, comprises all integer linear combinations of these vectors:

\begin{center}
    $L = \bigg\{\sum_{i=1}^{n} a_iv_i , a_i \in \Z \bigg\}$
\end{center}

If every $v_i$ ($i = 1, \ldots, n$) has integer coordinates, the lattice is termed an \textit{integral lattice}.

Figures \ref{fig:lattice0} and \ref{fig:lattice1} depict the lattice $L$ with bases $v=(3, 1)$ and $w=(-1, 1)$, and another basis configuration, respectively.

\begin{figure}[H]
    \begin{minipage}[b]{0.50\textwidth}
        \includegraphics[width=\textwidth]{./img/lattice_b0.png}
        \caption{Lattice $L$ spanned by $v, w$}
        \label{fig:lattice0}
    \end{minipage}
    \hspace{\fill}
    \begin{minipage}[b]{0.50\textwidth}
        \includegraphics[width=\textwidth]{./img/lattice_b1.png}
        \caption{Lattice $L$ spanned by $v', w'$}
        \label{fig:lattice1}
    \end{minipage}
\end{figure}

\section{Lattice Problems}

\subsection{SVP}

\textbf{The Shortest Vector Problem} (\texttt{SVP}): \textit{Find a nonzero vector $v \in L$ that minimizes the Euclidean norm $\lVert v \rVert$.}

Gauss developed an algorithm to determine an optimal basis for a two-dimensional lattice from any given basis. This method outputs the shortest nonzero vector in $L$, thereby solving the \texttt{SVP}.

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.6\textwidth]{./img/gauss_svp.png}
    \caption{Gauss reduction}
    \label{fig:gauss_svp}
\end{figure}

\begin{algorithm}[H]
    \vspace*{5px}
    \While{true}{
        \If{$\lVert v_2 \rVert < \lVert v_1 \rVert$}{
            swap $v_1, v_2$\;
        }
        m = $\left \lfloor{(v_1 \cdot v_2) {\lVert v_1 \rVert}^{-2}} \right \rfloor$\;
        \If{$m = 0$}{
            return $v_1, v_2$\;
        }
        $v_2 = v_2 - mv_1$\;
    }
    \caption{Gauss Basis Reduction}
\end{algorithm}

\textbf{Example}: In a lattice $L$ spanned by $v_1 = (8, 4), v_2 = (7, 5)$, applying the Gauss reduction algorithm results in $w_1 = (1, -1), w_2 = (6, 6)$ as shown in \ref{fig:gauss_svp}. $w_1$ represents the shortest nonzero vector in lattice $L$.

The complexity of solving \texttt{SVP} increases with the dimension of the lattice, making it a challenging problem without a known polynomial-time solution.

\subsection{CVP}

\textbf{The Closest Vector Problem} (\texttt{CVP}): \textit{Given a vector $t \in \R^m$ that is not in $L$, find the vector $v \in L$ closest to $t$, minimizing the Euclidean norm $\lVert t - v \rVert$.}

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.6\textwidth]{./img/cvp.png}
    \caption{CVP}
    \label{fig:cvp}
\end{figure}

\textbf{Example}: For a lattice $L$ spanned by $v_1 = (0, 1), v_2 = (2, 0)$, with target vector $t = (2.3, 2.4)$, the closest vector in the lattice $L$ is $w = (2, 2)$, as depicted in \ref{fig:cvp}.

Both \textbf{CVP} and \textbf{SVP} are considered $\mathcal{NP}$-hard problems, and their complexity increases with the dimension of the lattice. They serve as primitives for creating trapdoor functions in cryptographic schemes.

Several algorithms can approximate solutions for both \textbf{SVP} and \textbf{CVP}, such as \textbf{LLL}\cite{LLL} and \textbf{Babai's algorithm}\cite{babai} respectively.

\chapter{LLL}

\section{Introduction}

The \textbf{Lenstra-Lenstra-Lovász (LLL)} algorithm is a polynomial-time method designed to compute a "short" basis for a lattice.

\begin{theorem}
    \textbf{LLL}
\end{theorem}

Given a lattice $L \in \Z^n$ spanned by $B = \{v_1,\ldots,v_n\}$, the LLL algorithm produces a reduced basis $\{w_1, \ldots, w_n\}$ such that:

\begin{center}
    $\lVert w_i \rVert \le 2^{\frac{n(n-1)}{4(n-i+1)}} \cdot det(L)^{\frac{1}{n-i+1}}$ for $i=1,\ldots,n$
\end{center}

This process is executed in time polynomial in $n$ and the bit-size of the entries in the basis matrix $B$.\\

In this reduced basis, the vectors are as short as possible starting from the first, with each subsequent vector being longer. Additionally, these vectors are almost orthogonal to each other, meaning the dot product $w_i \cdot w_j$ is close to zero.

\subsection*{Example}

Consider the following basis, represented in matrix form, that spans a lattice $L$:

\[
L = 
\begin{pmatrix}
    4 & 9 & 10\\
    2 & 1 & 30\\ 
    3 & 7 & 9
\end{pmatrix}
\]

After applying the LLL algorithm, the basis is transformed to:

\[
LLL(L) = 
\begin{pmatrix}
    -1 & -2 & -1\\
     3 & -2 &  1\\
    -1 & -1 &  5
\end{pmatrix}
\]

Here, the first row is the shortest vector in the lattice $L$, thus addressing the \textbf{SVP} problem. For higher dimensions, however, the LLL algorithm provides only an approximation to the \textbf{SVP}.

\section{Algorithm}

\begin{algorithm}[H]
    \vspace*{5px}
    Input basis ${v_1, \ldots, v_n}$ \; \\
    $k = 2$ \; \\
    $v_1* = v_1$ \; \\
    \While{$k \le n$}{
        \For{$j = k -1$ \KwTo $1$ step $-1$} {
            $v_k = v_k - \left\lfloor\mu_{k, j}\right\rceil \boldsymbol{v}_{j}$  [size-reduction] \;
        }
        \eIf{$\lVert v_k^* \rVert^2 \ge (\frac{3}{4} - \mu_{k, k-1}^2) \lVert v_{k-1}^2 \rVert$}{
            $k = k + 1$ \;  [Lovàsz-condition]
        }{
            Swap $v_{k-1}$ and $v_k$ [swap-step] \; \\
            $k = \max(k - 1, 2)$
        }
    }
    Note: At each step, $v_1^*, \ldots, v_k^*$ is the orthogonalized set of vectors
    obtained with the Gram-Schmidt \ref{alg:gram_schmidt}. \\
    \vspace*{5px}
    $\mu_{i,j}$ is the quantity $(v_i \cdot v_j^*) / \lVert v_j^* \rVert^2$. \\
    Output reduced basis ${v_1, \ldots, v_n}$
    \caption{LLL reduction algorithm}
\end{algorithm}

The running time of the algorithm is polynomial and completes the main loop in no more than
$\mathcal{O}(n^2\log n + n^2\log M)$ steps, where $M = max \lVert v_i \rVert$. \\

Kelby Ludwig elucidates the intuition behind each step of the algorithm \cite{lllintuition}, and Oded Regev provides a detailed explanation of its properties \cite{lllexplanation}.\\

Applications of the LLL algorithm include:

\begin{enumerate}
    \item Factoring polynomials over the integers, e.g., factorizing $x^2 - 1$ into $(x + 1)(x - 1)$.
    \item Solving Integer Programming, a renowned $\mathcal{NP}$-complete problem, in polynomial time when the number of variables is fixed.
    \item Approximating solutions to the \textbf{CVP} and \textbf{SVP}, among other lattice-based challenges.
    \item Cryptanalysis, specifically breaking cryptographic protocols.
\end{enumerate}

% \chapter{LLL}
% 
% \section{Introduction}
% 
% The \textbf{Lenstra-Lenstra-Lovász} \textit{LLL}\cite{LLL} or \textit{$L^3$} is a polynomial time algorithm to find a ``short'' basis of a lattice.
% 
% \begin{theorem}
%     \textbf{LLL}
% \end{theorem}
% 
% \textit{Let $L \in \Z^n$ be a lattice spanned by $B = \{v_1,\ldots,v_n\}$. The LLL algorithm outputs a reduced lattice
% basis $\{w_1, \ldots, w_n\}$ with}
% 
% \begin{center}
%     $\lVert w_i \rVert \le 2^{\frac{n(n-1)}{4(n-i+1)}} det(L)^{\frac{1}{n-i+1}}$ for $i=1,\ldots,n$
% \end{center}
% 
% \textit{in time polynomial in n and in the bit-size of the entries of the basis matrix $B$}.\\
% 
% Basically the first vector of the new basis will be as short as possible, and the other will have increasing lengths. The new vectors
% will be as orthogonal as possible to one another, i.e., the dot product $w_i \dd w_j$ will be close to zero.
% 
% \subsection*{Example}
% 
% For example we can take the following basis (the rows are the vector) that span
% a lattice $L$.
% 
% \[
% L = 
% \begin{pmatrix}
%     4 & 9 & 10\\
%     2 & 1 & 30\\ 
%     3 & 7 & 9
% \end{pmatrix}
% \] 
% 
% Applying the LLL algorithm we obtain
% 
% \[
% LLL(L) = 
% \begin{pmatrix}
%     -1 & -2 & -1\\
%      3 & -2 &  1\\
%     -1 & -1 &  5
% \end{pmatrix}
% \] 
% 
% Where the first row is the shortest vector in the lattice $L$, and so solves the \textbf{SVP} problem.
% For higher dimensions however the LLL algorithm outputs only an approximation for the \textbf{SVP} problem.
% 
% \section{Algorithm}
% 
% \begin{algorithm}[H]
%     \vspace*{5px}
%     Input basis ${v_1, \ldots, v_n}$ \; \\
%     $k = 2$ \; \\
%     $v_1* = v_1$ \; \\
%     \While{$k \le n$}{
%         \For{$j = k -1$ \KwTo $1$ step $-1$} {
%             $v_k = v_k - \left\lfloor\mu_{k, j}\right\rceil \boldsymbol{v}_{j}$  [size-reduction] \;
%         }
%         \eIf{$\lVert v_k^* \rVert^2 \ge (\frac{3}{4} - \mu_{k, k-1}^2) \lVert v_{k-1}^2 \rVert$}{
%             $k = k + 1$ \;  [Lovàsz-condition]
%         }{
%             Swap $v_{k-1}$ and $v_k$ [swap-step] \; \\
%             $k = \max(k - 1, 2)$
%         }
%     }
%     Note: At each step, $v_1^*, \ldots, v_k^*$ is the orthogonalized set of vectors
%     obtained with the Gram-Schmidt \ref{alg:gram_schmidt}. \\
%     \vspace*{5px}
%     $\mu_{i,j}$ is the quantity $(v_i \cdots v_j^*) / \lVert v_j^* \rVert^2$. \\
%     Output reduced basis ${v_1, \ldots, v_n}$
%     \caption{LLL reduction algorithm}
% \end{algorithm}
% 
% The running time of the algorithm is polynomial and executes the main loop no more than
% $\mathcal{O}(n^2\log n + n^2\log M)$ steps, where $M = max \lVert v_i \rVert$. \\
% 
% Kelby Ludwig explains an intuition behind every step of the algorithm \cite{lllintuition}, and Oded Regev
% gives a more rigorous explanation of the algorithm and the properties \cite{lllexplanation}.\\
% 
% Some applications of the algorithm are the following:
% 
% \begin{enumerate}
%     \item Factoring polynomials over the integers. For example, given $x^2 - 1$ factor it into $x + 1$ and $x - 1$.
%     \item Integer Programming. This is a well-known $\mathcal{NP}$-complete problem. Using \textbf{LLL}, one can obtain a polynomial time solution
%           to integer programming with a fixed number of variables.
%     \item Approximation to the \textbf{CVP} or \textbf{SVP}, as well as other lattice problems.
%     \item Cryptanalysis (break cryptographic protocols).
% \end{enumerate}

\chapter{RSA Cryptanalysis}

\section{RSA Introduction}

\subsection{Algorithm}

\textbf{RSA} is one of the earliest and most used asymmetric cryptosystem.
The steps to generate a public/private key for \textbf{RSA} are:

\begin{enumerate}
    \item Fix $e = 65537$ or $e = 3$ (public).
    \item Find two primes $p, q$ such that $p - 1$ and $q - 1$ are relatively prime to $e$, gcd$(e, p-1) = 1$ and gcd$(e, q-1) = 1$.
    \item Compute $N = p * q$ and $\phi(n) = (p-1) * (q-1)$
    \item Calculate $d$ (private) as the multiplicative inverse of $e$ modulo $\phi(n)$.
    \item $(N, e)$ is the public key, $(N, d)$ is the private key.
\end{enumerate}

To encrypt a message $m$ with \textbf{textbook RSA}:

\begin{center}
    $c = m^e \mod N$
\end{center}

To decrypt a ciphertext $c$:

\begin{center}
    $m = c^d \mod N$
\end{center}

\subsection{Security}

RSA relies on the hardness of factoring $N$, and there isn't a polynomial algorithm for factorization, so it's considered secure.
However, there are different attacks against textbook RSA or bad implementations:

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ (the modulo operation is not applied), and we only need to find the $e$th root of $c$ over the
        integers to find $m$. This is one of the reasons why we pad the message before the encryption.
    \item If $q = p + x$ where $x$ is small enough, then it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and 
            $p = A - x$ for different values of $x$ until $N \mod p = 0$, which means that $q = N / p$.
    \item It's possible to recover $d$ if it is too small (in practice it could be used to accelerate decryption), using the Wiener's attack \cite{wiener}.
    \item Other attacks can be found on the excellent survey by Dan Boneh \cite{boneh99twentyyears}.
\end{itemize}

\section{Lattices against RSA}

We begin by discussing the main concepts behind Coppersmith's attack \cite{coppersmith96} and the underlying mathematics. This attack is applicable to \textbf{RSA} when a small exponent $e$ is used, and it can also be used to factorize the modulus if partial information about $p$ or $q$ is known. In 2017, this method was utilized to exploit \href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361}, also known as \textbf{ROCA} (Return Of Coppersmith's Attack) \cite{roca}.

\subsection{Mathematical Introduction}

Finding roots of a univariate polynomial over the integers is straightforward. However, finding roots of a modular polynomial, such as:

\begin{center}
    $f(x) \equiv 0 \mod N$
\end{center}

poses significant challenges. Assume $N$ is an \textbf{RSA} modulus whose factorization is unknown. Consider a univariate integer polynomial $f(x)$ of degree $n$:

\begin{center}
    $f(x) = x^n + a_{n-1}x^{n-1} + a_{n-2}x^{n-2} + \cdots + a_1n + a_0$
\end{center}

Coppersmith demonstrated that it is possible to recover the value $x_0$ such that $f(x_0) \equiv 0 \mod N$, with $x_0 < N^{\frac{1}{n}}$, in polynomial time using the following theorem:

\clearpage

\begin{theorem}
    \textbf{Howgrave-Graham}
\end{theorem}

\textit{Let $g(x)$ be a univariate polynomial with $n$ monomials and $m$ be a positive integer. If we establish a constraint $X$ and the following conditions are satisfied:}

\begin{center}
    \begin{eqnarray}
        g(x_0) \equiv 0 \mod N^m, |x_0| \le X \label{eq:bound} \\
         \lVert g(xX) \rVert < \frac{N^m}{\sqrt{n}} \label{eq:pol}
    \end{eqnarray}
\end{center}

\textit{Then $g(x_0) = 0$ is also valid over the integers.}

This theorem enables the computation of the root of $f(x) \mod N$ by finding a polynomial $g(x)$ that shares the same root, but modulo $N^m$. If the conditions \ref{eq:bound} and \ref{eq:pol} are met, the root of $g(x)$ over the integers corresponds to the same root $x_0$ such that $f(x_0) \equiv 0 \mod N$.

The idea behind Howgrave-Graham's method involves generating polynomials $p_i$ that also have $x_0$ as a root modulo $N^m$.

The significance of the \textbf{LLL} algorithm in this context is two-fold:

\begin{itemize}
    \item It conducts \textbf{integer linear operations} on basis vectors, ensuring that even if the basis changes, it remains a linear combination of vectors retaining $x_0$ as a root modulo $N^m$.
    \item If the lattice is properly constructed, the norm of the shortest vector in the reduced basis will satisfy \ref{eq:pol}, allowing the algorithm to predict the length's bound of the shortest vector it can find.
\end{itemize}

Polynomials $p_i$ ($g_{i,j}$ and $h_i$) sharing the same root $x_0$ over $N^m$ of $f$, where $\delta$ is the degree of $f$, can be crafted as follows:

\begin{center}
    \begin{eqnarray}
        g_{i,j}(x) &= x^j \cdot N^i \cdot f^{m-i}(x) \text{ for } i = 0,\hdots,m-1,\hspace{2mm} j=0,\hdots,\delta-1 \label{eq:create_pol} \\
        h_i(x) &= x^i \cdot f^m(x) \text{ for } i = 0,\hdots,t-1 
    \end{eqnarray}
\end{center}

Applying \textbf{LLL} to a lattice built from the coefficients of \ref{eq:create_pol}, we'll find a short vector $v = g(xX)$ that will satisfy \ref{eq:pol}. Consequently, if we then compute $g(x)$, we will be able to determine the root over the integers.

% \section{Lattices against RSA}
% 
% We start introducing the main ideas of Coppersmith's attack \cite{coppersmith96} and the math behind it.
% The attack can be used against \textbf{RSA} when a small $e$ is used, but also
% to factorize the modulus if some bits of $p$ or $q$ are known. In 2017 the attack was used to exploit
% \href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361} also named \textbf{ROCA}
% (Return Of Coppersmith's Attack) \cite{roca}.
% 
% \subsection{Mathematical introduction}
% 
% It's easy to find the roots of a univariate polynomial over the integers. Finding the roots of \textbf{modular} polynomial is hard, example:
% 
% \begin{center}
%     $f(x) \equiv 0 \mod N$
% \end{center}
% 
% Suppose $N$ is an \textbf{RSA} modulus, and we don't know the factorization of it. Let's have an univariate integer polynomial $f(x)$ with degree $n$
% 
% \begin{center}
%     $f(x) = x^n + a_{n-1}x^{n-1} + a_{n-2}x^{n-2} + \cdots + a_1n + a_0$
% \end{center}
% 
% Coppersmith showed how we can recover the value $x_0$ such that $f(x_0) \equiv 0 \mod N$, with $x_0 < N^{\frac{1}{n}}$ in polynomial time
% using the following theorem
% 
% \clearpage
% 
% \begin{theorem}
%     \textbf{Howgrave-Graham}
% \end{theorem}
% 
% \textit{Let $g(x)$ be an univariate polynomial with n monomials and $m$ be a positive integer.
%     If we have some restraint $X$ and the following equations hold}
% 
% \begin{center}
%     \begin{eqnarray}
%         g(x_0) \equiv 0 \mod N^m, |x_0| \le X \label{eq:bound} \\
%          \lVert g(xX) \rVert < \frac{N^m}{\sqrt{n}} \label{eq:pol}
%     \end{eqnarray}
% \end{center}
% 
% \textit{Then $g(x_0) = 0$ holds over the integers.}
% 
% \vspace*{10px}
% 
% This theorem states that is possible to compute the root of $f(x) \mod N$ if we can find a polynomial $g(x)$ that share the same root but modulo $N^m$.
% If \ref{eq:bound} and \ref{eq:pol} hold then we can simply compute the root of $g(x)$ over the integers to have the same root $x_0$
% such that $f(x_0) \equiv 0 \mod N$.
% 
% Howgrave-Graham's idea is to find this polynomial $g$ by combining polynomials $p_i$ who also have $x_0$ as roots modulo $N^m$.
% 
% \vspace*{10px}
% 
% The \textbf{LLL} algorithm is fundamental because:
% 
% \begin{itemize}
%     \item It only does \textbf{integer linear operations} on the basis vectors.
%         In this way even if the basis is different it's only a linear combination of vector that still have $x_0$ as root
%         modulo $N^m$.
%     \item If we craft the lattice properly, the norm of the shortest vector on the reduced basis will satisfy \ref{eq:pol}.
%         We know the length's bound of the shortest vector that LLL could find.
% \end{itemize}
% 
% We can easily create polynomials $p_i$ ($g_{i,j}$ and $h_i$)  sharing the same root $x_0$ over $N^m$ of $f$ where $\delta$ is the degree of $f$:
% 
% \begin{center}
%     \begin{eqnarray}
%         g_{i,j}(x) &= x^j \cdot N^i \cdot f^{m-i}(x) \text{ for } i = 0,\hdots,m-1,\hspace{2mm} j=0,\hdots,\delta-1 \label{eq:create_pol} \\
%         h_i(x) &= x^i \cdot f^m(x) \text{ for } i = 0,\hdots,t-1 
%     \end{eqnarray}
% \end{center}
% 
% Applying \textbf{LLL} to a lattice constructed by the coefficients of \ref{eq:create_pol} we'll find a short vector $v = g(xX)$ that will
% satisfy \ref{eq:pol}. If we then take $g(x)$ we will be able to compute the root over the integer.

\subsection{Recovering Plaintext with Known MSB Bits of $m$}

We will employ a simplified version of Coppersmith's attack, effective when the root is approximately less than $N^{\frac{1}{6}}$.

\textbf{Problem Setup}. Consider an \textbf{RSA} modulus $N$ of 97 bits:

\begin{center}
    $N=\texttt{0x1a34992f64135aedaa0fe2bfd}$ and $e = 3$.
\end{center}

Before encryption, the message $m$ undergoes padding:

\[
    z = pad || m = \texttt{0xffffffffffffffffffffff} || m
\]

where $||$ denotes concatenation.

The resulting ciphertext is:

\[
    c = z^e \mod N = \texttt{0x11701d5479bcdef207cba6937}
\]

Assuming the factorization of $N$ is unknown, our goal is to deduce the message $m$. We are aware of the padding and that the message $m$ is 2 bytes long, consisting only of ASCII characters, so $m < \texttt{0x7b7b}$.

\vspace*{10px}

Let's define:

\[
    a = \texttt{0xffffffffffffffffffffff0000}
\]

as the known padded string that was encrypted.

\vspace*{10px}

This leads to the equation $c = (a + m)^3 \mod N$, for an unknown small $m$. We can then define $f(x) = (a + x)^3 - c$ and set up the problem to find a small root $m$ such that $f(m) \equiv 0 \mod N$:

\begin{align*}
    f(x) = x^3 + \texttt{0x17d7f55d0c9dc5851af54957c}x^2 + \texttt{0x3cf96fed2ea927527a2ed329}x\\
    + \texttt{0x32882e547964b3bcf75d315e}
\end{align*}

\vspace*{10px}

\textbf{Lattice Construction}. Let the coefficients of $f(x)$ be $f(x) = x^3 + f_2x^2 + f_1x + f_0$ and $X = \texttt{0x7b7b}$ be the upper bound for the root $m$. We can construct the matrix:

\[
B = 
\label{mat:lattice_rsa}
\begin{pmatrix}
    X^3 & f_2X^2 & f_1X & f_0 \\
    0 & NX^2 & 0 & 0 \\
    0 & 0 & NX & 0 \\
    0 & 0 & 0 & N
\end{pmatrix}
\]

Each row of this matrix corresponds to the coefficient vectors of the polynomials $f(x), Nx^2, Nx,$ and $N$, all evaluated at $x = m$ being 0 modulo $N$.
We apply Howgrave-Graham with $m=1$ (the $N^m$ parameter, not the message).

In this lattice setup, every vector is of the form $v = (v_3X^3, v_2X^2, v_1X, v_0)$, because any integer linear combination of the vectors in the lattice will respect the bound $X^i$ for $i=\dim(B)-1,..,0$.

\textbf{Apply LLL}. We apply LLL to identify the shortest vector in the reduced basis:

\[
    \begin{split}
    v = (-\texttt{0xd55d67baf71d}X^3, \texttt{0x3cf3cca3200a58bf}X^2, \\
    \texttt{0x3854d44211e80f248449}X, -\texttt{0xec93bf51cf766f7b9f1e5e6})
    \end{split}
\]

Using the coefficients of $v$, we construct the polynomial $g$:

\[
    \begin{split}
        g(x) = -\texttt{0xd55d67baf71d}x^3 +  \texttt{0x3cf3cca3200a58bf}x^2 \\
        \texttt{0x3854d44211e80f248449}x -\texttt{0xec93bf51cf766f7b9f1e5e6}
    \end{split}
\]

% \subsection{Recover plaintext with known MSB bits of $m$}
% 
% We will use a simplified version of the Coppersmith's attack that works when the root is approximately less than $N^{\frac{1}{6}}$.
% 
% \textbf{Problem setup}. Let $N$ be an \textbf{RSA} modulus of $97$-bit:
% 
% \begin{center}
%     $N=\texttt{0x1a34992f64135aedaa0fe2bfd}$ and $e = 3$.
% \end{center}
% 
% Before the encryption the message $m$ is padded as
% 
% \[
%     z = pad || m = \texttt{0xffffffffffffffffffffff} || m
% \]
% 
% where $||$ is the concatenation.
% 
% The ciphertext is
% 
% \[
%     c = z^e \mod N = \texttt{0x11701d5479bcdef207cba6937}
% \]
% 
% Suppose that we don't know the factorization of $N$, and we would like to know the message $m$.
% We know the padding and that the message $m$ is 2 bytes long, composed by only ascii letters, so $m < \texttt{0x7b7b}$.
% 
% \vspace*{10px}
% 
% Let's define
% 
% \[
%     a = \texttt{0xffffffffffffffffffffff0000}
% .\] 
% 
% which is the known padding string that got encrypted.
% 
% \vspace*{10px}
% 
% Thus, we have that $c = (a + m)^3 \mod N$, for an unkown small $m$.
% We can define $f(x) = (a + x)^3 - c$, and setup the problem to find a small root $m$ such that $f(m) \equiv 0 \mod N$
% 
% \begin{align*}
%     f(x) = x^3 + \texttt{0x17d7f55d0c9dc5851af54957c}x^2 + \texttt{0x3cf96fed2ea927527a2ed329}x\\
%     + \texttt{0x32882e547964b3bcf75d315e}
% \end{align*}
% 
% \vspace*{10px}
% 
% \textbf{Lattice contruction}. Let the coefficients of $f$ be $f(x) = x^3 + f_2x^2 + f_1x + f_0$
% and $X = \texttt{0x7b7b}$ be the upper bound of the size of the root $m$. We can construct the matrix
% 
% \[
% B = 
% \label{mat:lattice_rsa}
% \begin{pmatrix}
%     X^3 & f_2X^2 & f_1X & f_0 \\
%     0 & NX^2 & 0 & 0 \\
%     0 & 0 & NX & 0 \\
%     0 & 0 & 0 & N
% \end{pmatrix}
% \] 
% 
% The rows of the matrix correspond to the coefficient vectors of the polynomials $f(x), Nx^2, Nx$ and $N$, furthermore we know
% that each polynomial will be 0 modulo $N$ if evaluated at $x = m$.
% We applied Howgrave-Graham with $m=1$ (the $N^m$ parameter not the message).\\
% With this lattice construction every vector is of the form $v = (v_3X^3, v_2X^2, v_1X, v_0)$, because any
% integer linear combination of the vector of the lattice will keep the bound $X^i$ for $i=\dim(B)-1,..,0$.
% 
% \textbf{Apply LLL}. We then apply LLL to find the shortest vector of the reduced basis:
% 
% \[
%     \begin{split}
%     v = (-\texttt{0xd55d67baf71d}X^3, \texttt{0x3cf3cca3200a58bf}X^2, \\
%     \texttt{0x3854d44211e80f248449}X, -\texttt{0xec93bf51cf766f7b9f1e5e6})
%     \end{split}
% \]
% 
% We can construct the polynomial $g$ using the coefficients of $v$
% 
% \[
%     \begin{split}
%         g(x) = -\texttt{0xd55d67baf71d}x^3 +  \texttt{0x3cf3cca3200a58bf}x^2 \\
%         \texttt{0x3854d44211e80f248449}x -\texttt{0xec93bf51cf766f7b9f1e5e6}
%     \end{split}
% \]

Knowing that:

\[
    g(x_0) \equiv 0 \mod N, |x_0| \le X
\]

We need to verify that:

\[
    \lVert g(xX) \rVert \le \frac{N}{\sqrt{n}}
\]

In this scenario, the determinant of $B$, $\text{det} B = X^6N^3$, and LLL is expected to find a short vector with $\lVert v \rVert \le 2^{\frac{n(n-1)}{4(n)}} (\det B)^{\frac{1}{n}}$.
Neglecting the factor $2^{\frac{3}{4}}$ (given that $n = 4$), the requirement simplifies to:

\[
    g(m) \le \lVert v \rVert \le (\det B)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}
\]

\vspace*{10px}

We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}$. Solving for $X$ under this constraint implies $X < (\frac{N}{16})^{\frac{1}{6}}$. Adjusting for the $\sqrt{4}$ as an error term allows $X < N^{\frac{1}{6}}$.
This inequality holds with the given numbers; however, even if the bound of the shortest vector is larger, there remains a chance to find the correct root.

Computing the root of $g(x)$ over the integers, we find $m = \texttt{0x4142} = $ ``AB'', which is the correct result.\\

This specific lattice construction is effective for finding roots up to size $N^{\frac{1}{6}}$. Therefore, the same approach could be applied to recover:

\begin{itemize}
    \item Approximately 170 unknown bits of a message from an RSA 1024-bit modulus.
    \item Approximately 341 unknown bits of a message from an RSA 2048-bit modulus.
    \item Approximately 683 unknown bits of a message from an RSA 4096-bit modulus.
\end{itemize}

For larger roots, a more extensive lattice incorporating additional polynomials generated with \ref{eq:create_pol} is required.
This methodology is further elaborated in \cite{may2011}, with a detailed implementation in SageMath by David Wong \cite{wong2015}.

\subsection{Factorizing the Modulus with Known MSB Bits of $p$}

Lattices can also assist in factorizing a modulus $N$ if the most significant bits of one of its factors are known. Let $a = 2^lb$, where $b$ represents the known bits of $p$, then $p = a + r$ for some small $r$.
We can reformulate this to find the root $r$ of $f(x) = a + x \mod p$ such that $f(r) = p \equiv 0 \mod p$.
Assume we know $r$'s upper bound, $|r| < X$.

We can configure the following lattice:

\[
    B = 
    \begin{pmatrix}
        X^2 & Xa & 0 \\
        0 & X & a \\
        0 & 0 & N
    \end{pmatrix} 
\]

\vspace*{10px}

Each row corresponds to the polynomials $x(x + a)$, $(x + a)$, and $N$, each evaluated at $x = r$ being $0 \mod p$. Similar to \ref{mat:lattice_rsa}, every polynomial is scaled by $X$, the upper bound.

The shortest vector found by \textbf{LLL} should be less than $p$ because if $g(x) < p$, then by construction $g(r) = 0$ over the integers can be computed. We calculate:

\[
    (detB)^{\frac{1}{\dim L}} = (X^3N)^{\frac{1}{3}} < p
\]

Solving for $X$ gives $X < p^{\frac{1}{3}}$, indicating that this method is valid for every $|r| < p^{\frac{1}{3}}$.

To pinpoint $r$, simply apply \textbf{LLL} and take the first vector (the shortest), which will have the form $v = (v_2X^2, v_1X, v_0)$.

Finally, construct $g(x) = v_2x^2 + v_1x + v_0$ (adjust each coefficient by dividing by $X^i$) and compute the roots over the integers to find $r$. We executed this attack against an \textbf{RSA}-$1024$ modulus knowing the $160$ most significant bits of $p$. This strategy can be extended to $X < p^{\frac{1}{2}}$ with a larger lattice.

% We know that
% 
% \[
%     g(x_0) \equiv 0 \mod N, |x_0| \le X
% \]
% 
% What we need to prove is that
% 
% \[
%     \lVert g(xX) \rVert \le \frac{N}{\sqrt{n}}
% \] 
% 
% In this example, det $B = X^6N^3$, and LLL will find a short vector with $\lVert v \rVert \le 2^{\frac{n(n-1)}{4(n)}} (\det B)^{\frac{1}{n}}$.
% If we ignore the $2^{\frac{3}{4}}$ factor (remember that $n = 4$), then we need to satisfy
% 
% \[
%     g(m) \le \lVert v \rVert \le (\det B)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}
% \] 
% 
% \vspace*{10px}
% 
% We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}$, if we solve for $X$ this will be satisfied when
% $X < (\frac{N}{16})^{\frac{1}{6}}$. We can consider the $\sqrt{4}$ an error term to have $X < N^{\frac{1}{6}}$.
% With the numbers we have this inequality is verified, however even if the bound of the shortest vector is larger we still have some possibilities to
% find the correct root.
% 
% Computing the root of $g(x)$ over the integers we obtain $m = \texttt{0x4142} = $ ``AB'' which is the correct result.\\
% 
% This specific lattice works to find roots up to size $N^{\frac{1}{6}}$, so the same construction will work if we want to find
% 
% \begin{itemize}
%     \item $\approx$ 170 unkown bits of message from an RSA 1024-bit modulus
%     \item $\approx$ 341 unkown bits of message from an RSA 2048-bit modulus
%     \item $\approx$ 683 unkown bits of message from an RSA 4096-bit modulus
% \end{itemize}
% 
% To compute bigger root a bigger lattice with more polynomials generated with \ref{eq:create_pol} is needed,
% this method is better described in \cite{may2011} and a detailed implementation in sagemath by David Wong is available \cite{wong2015}.
% 
% \subsection{Factorize modulus with known MSB bits of $p$}
% 
% Lattices can also be used to factor a modulus $N$ knowing the most significant bits of one of the factor.
% Let $a = 2^lb$ where $b$ are the bits known of $p$, then $p = a + r$ for some small value of $r$.
% We can rewrite the problem to find the root $r$ of $f(x) = a + x \mod p$ such that $f(r) = p \equiv 0 \mod p$.
% Suppose that we know the bound of $r$ such that $|r| < X$ for some bound $X$.
% 
% We can construct the following lattices
% 
% \[
%     B = 
%     \begin{pmatrix}
%         X^2 & Xa & 0 \\
%         0 & X & a \\
%         0 & 0 & N
%     \end{pmatrix} 
% \]
% 
% \vspace*{10px}
% 
% Notice that the rows correspond to the polynomials $x(x + a)$, $(x + a)$, $N$ and that each of these polynomials will
% be $0 \mod p$ when $x = r$. As before \ref{mat:lattice_rsa}, every polynomial is scaled by $X$ which is the upperbound.
% 
% We need to verify that the shortest vector that we can find with \textbf{LLL} will be less than $p$ because if $g(x) < p$,
% then by construction we can compute $g(r) = 0$ over the integers.
% In this case we have
% 
% \[
%     (detB)^{\frac{1}{\dim L}} = (X^3N)^{\frac{1}{3}} < p
% \]
% 
% Solving for $X$ we obtain $X < p^{\frac{1}{3}}$, so this method works for every $|r| < p^{\frac{1}{3}}$.
% 
% To find $r$ we just need to apply \textbf{LLL} and take the first vector (shortest) that will be of the form $v = (v_2X^2, v_1X, v_0)$.
% 
% The last step is to construct $g(x) = v_2x^2 + v_1x + v_0$ (just divides every coefficient by $X^i$)
% and computes the roots over the integers to find $r$. We implemented the attack against a modulus
% of \textbf{RSA}-$1024$ knowing the $160$ most significant bits of $p$. The attack can be expanded
% until $X < p^{\frac{1}{2}}$ with a bigger lattice.

\subsection{Real World Case Study: ROCA}

\textbf{ROCA}\cite{roca} is a vulnerability discovered in 2017 in a software library \textit{RSALib}, which affects the generation of the public modulus for \textbf{RSA}. This issue is believed to impact millions of smart cards. The vulnerability arises from the method used to generate prime numbers; since embedded devices lack extensive computational resources, a rapid generation method was necessary. Each prime number generated adheres to the following structure:

\[
    p = kM + (g^a \mod M)
    \label{eq:roca}
\]

Here, $k, a$ are random values unique for each prime (unknown), $g = 65537$, and $M$ is a \textit{primorial} number. A \textit{primorial} number $M$ is defined as the product of the first $n$ successive primes:

\[
    M = P_n\# = \prod_{i=1}^{n} P_i
\]

where $n$ corresponds to the key size. In \textit{RSALib}, the primorial number used is of similar size to $p$, and the values of $a$ and $k$ are low in entropy. The modulus is expressed as:

\[
    N = (kM + g^{a}\mod M)(lM + g^{b}\mod M)
\]

for $a, b, k, l \in \Z$. This can be simplified to:

\[
    N \equiv g^{a + b} \equiv g^{c} \mod M
\]

for some $c \in \Z$, as all multiples of $kM$ or $lM$ reduce to $0 \mod M$.

Computing the discrete logarithm of $g^c \mod M$ is feasible due to the smoothness of $|M|$ and the effectiveness of the Pohlig-Hellman algorithm \cite{pohlig}. With $c$, we can establish a boundary for the value of $a$ as follows:

\[
    \frac{c}{2} \le a \le \frac{c + ord}{2}
\]

where $ord$ is the multiplicative order of the generator $g$ within $Z_M$.

Once the bounds for $a$ are known, it becomes possible to determine the value of $k$ using Coppersmith's method \cite{coppersmith96} for each plausible $a$.

The conceptual flow of the algorithm is outlined below:

\begin{algorithm}[H]
    \vspace*{5px}
    \textbf{Input: } $N, M, g, ord, X$ \; \\
    $c = \log_{g}N \mod M$ \;\\
    \For{$a \gets (\frac{c}{2}, \frac{c + ord}{2})$}{
        \vspace*{5px}
        $f(x) = xM + (g^a \mod M)$ \; \\
        $g(x) = \frac{1}{M}f(x)$ [make monic, the roots are the same] \; \\
        $k = $ coppersmith($g(x), N, X$) [find k] \; \\ 
        $p = kM + (g^a \mod M)$  [candidate] \; \\
        \If{$N \mod p = 0$}{
            return $p$ \;
        }
    }
    \caption{Exploit idea}
\end{algorithm}

However, in practice, the bound of $a$ is too large, leading the authors to devise a method to transpose the problem to a smaller $M'$ since the bit size of $M$ is more than sufficient to make the attack viable with Coppersmith's algorithm. Their optimization finds a new $M'$ such that:

\begin{itemize}
    \item $(p, q)$ maintain the form specified in \ref{eq:roca}, hence $M'$ must be a divisor of $M$.
    \item Coppersmith's algorithm can determine $k'$ for the accurate guess of $a'$.
    \item A smaller order of $g$ in $\Z_{M'}$ is found to reduce the size of $a'$.
\end{itemize}

We successfully replicated the attack in SageMath against \textbf{RSA}-$512$, using $M'$ as specified by Bruno Produit \cite{rocaopt} in his thesis and implementing David Wong's version of Coppersmith's attack \cite{wong2015}.

% \subsection{Real world case study: ROCA}
% 
% \textbf{ROCA}\cite{roca} is a vulnerability discovered in 2017 in a software library \textit{RSALib} in the generation
% of the public modulus for \textbf{RSA}. The attack is believed to affect millions of smart cards. The vulnerbality
% resides in the generation of prime numbers, since embedded devices don't have lots of computational resources a fast way
% was needed. Every prime generated has the following structure:
% 
% \[
%     p = kM + (g^a \mod M)
%     \label{eq:roca}
% \]
% 
% Where $k, a$ are random values different for every prime (unkown), $g = 65537$ and $M$ it's a \textit{primorial} number.
% A \textit{primorial} number $M$ is the product of the first $n$ successive primes 
% 
% \[
%     M = P_n\# = \prod_{i=1}^{n} P_i
% \]
% 
% where $n$ is related to key size. In the case of \textit{RSALib} the primorial number used had similar dimension to $p$,
% and the value $a, k$ didn't have much entropy. The modulus is
% 
% \[
%     N = (kM + g^{a}\mod M)(lM + g^{b}\mod M)
% \]
% 
% for $a, b, k, l \in \Z$. It's equivalent to
% 
% \[
%     N \equiv g^{a + b} \equiv g^{c} \mod M
% \]
% 
% for some $c \in \Z$ since all the multiples of $kM$ or $lM$ are $0 \mod M$.
% 
% It's possible to compute the discrete logarithm of $g^c \mod M$ since the size of $|M|$ is smooth with
% Pohlig-Hellman algorithm \cite{pohlig}. With $c$ we can set the bound to the value of $a$ since
% 
% \[
%     \frac{c}{2} \le a \le \frac{c + ord}{2}
% \]
% 
% where $ord$ is the multiplicative order of the generator $g \in Z_M$.
% 
% Once we know the bound of $a$ we can try to find the value of $k$ with the Coppersmith's method \cite{coppersmith96} for every possible $a$.
% 
% The concept of the algorithm is the following:
% 
% \begin{algorithm}[H]
%     \vspace*{5px}
%     \textbf{Input: } $N, M, g, ord, X$ \; \\
%     $c = \log_{g}N \mod M$ \;\\
%     \For{$a \gets (\frac{c}{2}, \frac{c + ord}{2})$}{
%         \vspace*{5px}
%         $f(x) = xM + (g^a \mod M)$ \; \\
%         $g(x) = \frac{1}{M}f(x)$ [make monic, the roots are the same] \; \\
%         $k = $ coppersmith($g(x), N, X$) [find k] \; \\ 
%         $p = kM + (g^a \mod M)$  [candidate] \; \\
%         \If{$N \mod p = 0$}{
%             return $p$ \;
%         }
%     }
%     \caption{Exploit idea}
% \end{algorithm}
% 
% In practice however the bound of $a$ is too big, so the authors found a way to move the same problem to a smaller $M'$ since the
% bitsize of $M$ are more than sufficient for making the attack works with Coppersmith's algorithm. Their optimization
% find a new $M'$ such that
% 
% \begin{itemize}
%     \item $(p, q)$ are still of the form \ref{eq:roca}, so $M'$ must be a divisor of $M$.
%     \item Coppersmith's algorithm will find $k'$ for the correct guess of $a'$.
%     \item Find a small order of $g \in \Z_M'$ to make $a'$ smaller.
% \end{itemize}
% 
% We were able to reproduce the attack in sagemath against \textbf{RSA}-$512$, and we used as $M'$ the value
% that Bruno Produit \cite{rocaopt} made available in his thesis and David Wong's implementation of Coppersmith's attack \cite{wong2015}.

\chapter{ECDSA Cryptanalysis}

\section{ECDSA Introduction}

\subsection{Algorithm}

\textbf{ECDSA} is a variant of the Digital Signature Algorithm (\textbf{DSA}) that incorporates elliptic curve cryptography. It relies on three public parameters:

\begin{itemize}
    \item The elliptic curve \textbf{$E$}.
    \item The generator point \textbf{$G$}.
    \item The order of the generator \textbf{$n$}.
\end{itemize}

Additionally, a private key $d \in [1, n-1]$ and a corresponding public key $Q = dG$ are required. To \textbf{sign} a message $m$:

\begin{enumerate}
    \item Compute $h = $ MSB(HASH$(m))$, where HASH is a cryptographic hash function, such as SHA-256, and MSB extracts the $n$ most significant bits.
    \item Select a random integer $k \in [1, n-1]$ (nonce).
    \item Calculate $P = kG = (x_1, y_1)$ and set $r = x_1$; if $x_1 = 0$, repeat the steps.
    \item Compute $s = k^{-1}(h + dr) \mod n$; if $s = 0$, repeat the steps.
    \item The signature comprises $(r, s)$.
\end{enumerate}

\clearpage

To \textbf{verify} the signature:

\begin{enumerate}
    \item Compute $h = $ HASH$(m)$.
    \item Calculate $u_1 = hs^{-1} \mod n$ and $u_2 = rs^{-1} \mod n$.
    \item Compute $P = u_1G + u_2Q = (x_1, y_1)$; if $P = O$, the signature is invalid.
    \item If $r \equiv x_1 \mod n$, the signature is valid.
\end{enumerate}

\subsection{Security}

The security of \textbf{ECDSA} hinges on the discrete logarithm problem.
Given the points $Q$ and $G$ such that $Q = kG$, it is a challenging problem to discover $k$. However, similar to RSA, ECDSA is susceptible to various implementation attacks:

\begin{itemize}
    \item If the same $k$ is used for generating two different signatures, it is possible to recover the private key $d$. This vulnerability was notably exploited in the PlayStation 3.
    \item The private key $d$ can be compromised if the nonce $k$ is not produced by a cryptographically secure pseudo-random number generator.
    \item If a non-standard (custom) elliptic curve is used, the discrete logarithm problem may become tractable, potentially undermining the security of the system \cite{giuliani99attackson}, \cite{mov}, \cite{smartass}.
\end{itemize}

\section{Lattices against ECDSA}

The lattice-based attacks discussed here derive from the \textit{Hidden Number Problem (HNP)}, initially introduced by Boneh and Venkatesan \cite{hnp}. This concept was later extended to demonstrate that even partial information about the nonce $k$, specifically its most significant bits, can compromise the entire nonce \cite{ecdsa_nonce}. In 2020, researchers successfully extracted the most significant bits of \textit{DH} and \textit{DHE} secrets in \textbf{TLS $\le$ 1.2} and employed a lattice construction similar to the one described here to solve the \textit{HNP} and thereby breach the protocol \cite{cryptoeprint:2020:1151} (Raccoon Attack).

% \chapter{ECDSA Cryptanalysis}
% 
% \section{ECDSA Introduction}
% 
% \subsection{Algorithm}
% 
% \textbf{ECDSA} is a variant of the Digital Signature Algorithm (\textbf{DSA}) which uses elliptic curve cryptography.
% There are 3 public parameters:
% 
% \begin{itemize}
%     \item The elliptic curve \textbf{$E$}.
%     \item The generator point \textbf{$G$}.
%     \item The generator's order \textbf{$n$}.
% \end{itemize}
% 
% We also need to create a private key $d \in [1, n-1]$ and public key $Q = dG$. To \textbf{sign} a message $m$:
% 
% \begin{enumerate}
%     \item Compute $h = $ MSB(HASH$(m))$ where HASH is a cryptographic hash function, for example SHA-256, and MSB take the $n$ most significant bits.
%     \item Select a random integer $k \in [1, n-1]$ (nonce).
%     \item Calculate $P = kG = (x_1, y_1)$ and set $r = (x_1)$ if $x_1 \neq 0$, otherwise repeat the steps.
%     \item Compute $s = k^{-1}(h + dr) \mod n$, if $s = 0$ repeat the steps.
%     \item The signature is composed by $(r, s)$.
% \end{enumerate}
% 
% \clearpage
% 
% To \textbf{verify} the signature:
% 
% \begin{enumerate}
%     \item Compute $h = $ HASH$(m)$.
%     \item Calculate $u_1 = hs^{-1} \mod n$ and $u_2 = rs^{-1} \mod n$.
%     \item Compute $P = u_1G + u_2Q = (x_1, y_1)$, if $P = O$ the signature is invalid.
%     \item If $r \equiv x_1 \mod n$ then the signature is valid.
% \end{enumerate}
% 
% \subsection{Security}
% 
% The security of \textbf{ECDSA} depends on the discrete logarithm problem.
% Given the points $Q, G$ such that $Q = k*G$ it's considered an hard problem to find $k$.
% However, as RSA, there are lots of implementation attacks
% 
% \begin{itemize}
%     \item If the same $k$ is used to generate two different signatures it's possible to recover the secret key $d$. This was a real bug 
%         discovered in the Playstation 3.
%     \item It's possible to recover $d$ if the parameter $k$ is not generated with a cryptographically secure pseudo random generator.
%     \item If the elliptic curve used is not standardized (custom) then it's possible that the discrete logarithm is easily solvable
%         \cite{giuliani99attackson}, \cite{mov} \cite{smartass}.
% \end{itemize}
% 
% \section{Lattices against ECDSA}
% 
% The lattice-based attacks presented here are based on the \textit{HNP (Hidden Number Problem)} introduced
% by Boneh and Venkatesan \cite{hnp} that was in a second moment generalized to prove that the most significant bits
% of $k$ can potentially leak the entire nonce \cite{ecdsa_nonce}.
% In 2020 a group of researcher were able to leak the MSBs of \textit{DH} and \textit{DHE} secrets in \textbf{TLS $\le$ 1.2}
% and used a similar lattice constructions as the one we will describe to solve the \textit{HNP}
% and thus break the protocol \cite{cryptoeprint:2020:1151} (Raccoon Attack).

% \subsection{Recover $d$ from a pair of signatures with small $k_1, k_2$}
% 
% \textbf{Problem setup}.
% Let $p = \texttt{0xffffffffffffd21f}$ and let $E: y^2 = x^3 + 3$ be an elliptic curve over $\F_p$ with the generator
% $G = (\texttt{0xcc3b3d1a0c4938ef}, \\ \texttt{0x4ab35ff66f8194fa})$ of order $n = \texttt{0xfffffffefa23f437}$.
% 
% We have two signature:
% 
% \[
%     \begin{array}{c}
%         (r_1, s_1) = (\texttt{0x269fa43451c5ff3c}, \texttt{0x1184ec0a74d4be7c}) \\
%         \text{and hash } h_1 = \texttt{0xb526aef1a341cfe6}
%     \end{array}
% \]
% 
% \[
%     \begin{array}{c}
%         (r_2, s_2) = (\texttt{0xf77cda14f5bf50a2}, \texttt{0xcd1143ccc1516b02}) \\ 
%         \text{and hash } h_2 = \texttt{0x84768ddee659efea}
%     \end{array}
% \]
% 
% And we know that both of these signatures use 32-bit nonce $k$, note that $n$ is a 64-bit number. We can set up the problem as a system of equation
% 
% \[
%     \begin{array}{c}
%         s_1 \equiv \displaystyle\frac{h_1 + dr_1}{k_1} \mod n \\ \\ 
%         s_2 \equiv \displaystyle\frac{h_2 + dr_2}{k_2} \mod n
%     \end{array}
% \]
% 
% We don't know $d, k_1, k_2$, but we can write 
% 
% \[
%     d = \displaystyle\frac{s_1k_1 - h_1}{r_1}
% \]
% 
% and rewrite the equation in
% 
% \[
%     k_1 - s_1^{-1}s_2r_1^{-1}k_2 + s_1^{-1}r_1h_2r_2^{-1} - s_1^{-1}h_1 \equiv 0 \mod n
% \]
% 
% To simplify the equation we write $t = - s_1^{-1}s_2r_1^{-1}k_2$ and $u = s_1^{-1}r_1h_2r_2^{-1} - s_1^{-1}h_1$. In this way we have
% 
% \[
%     \begin{array}{c}
%         k_1 + tk_2 + u \equiv 0 \mod n \\
%         \text{or} \\
%         -k_1 = tk_2 + u - xn \text{   for some $x$}
%     \end{array}
% \]
% 
% We know that $|k_1|, |k_2| < K = 2^{32}$.
% 
% \vspace*{10px}
% 
% \textbf{Lattice construction}. We construct the lattice $B = (b_0, b_1, b_2)$:
% 
% \[
%     B = 
%     \begin{pmatrix}
%         n & 0 & 0 \\
%         t & 1 & 0 \\
%         u & 0 & K
%     \end{pmatrix}
% \]
% 
% Note that the vector $v = (-k_1, k_2, K)$ is in the lattice because
% 
% \[
%     v = -xb_0 + k_2b_1 + b_2 = (-xn + tk_2 + u = -k_1, k_2, K)
% \]
% 
% for some $x, k_2 \in Z$.
% 
% \vspace*{10px}
% 
% Can we prove that $v$ is a short vector that we can find with LLL?
% 
% We have 
% 
% \begin{center}
%     $\lVert v \rVert = \sqrt{k_1^2 + k_2^2 + K^2} \le \sqrt{3K^2} = \sqrt{3}K $
% \end{center}
% 
% And we expect the shortest vector to have length
% 
% \[
%     \begin{array}{c}
%         \approx 2^{\frac{1}{2}}(\det B)^{\frac{1}{3}} \\
%         \approx 2^{\frac{1}{2}}(nK)^{\frac{1}{3}}
%     \end{array}
% \]
% 
% So we want that 
% 
% \begin{center}
%     $\lVert v \rVert \le \sqrt{3}K < \sqrt{2}(nK)^{\frac{1}{3}}$
% \end{center}
% 
% And if we remove the smaller terms this will be satisfied when $K < (nK)^{\frac{1}{3}}$ or $K < \sqrt{n}$.
% 
% In this case if we apply LLL in the second row we obtain:
% 
% \begin{center}
%     $v = (-k_1, k_2, K) = (\texttt{-0x50a65330, 0x1f5b977a, 0x100000000})$
% \end{center}
% 
% To retrieve $d$ we just need to compute 
% 
% \[
%     d = r_1^{-1}(k_1s_1 - h_1) = \texttt{0xf00e5fb275bfd304}
% \]

%Boneh-Venkatesan1996_Chapter_HardnessOfComputingTheMostSign-1.pdf, HPL-1999-90-1.pdf, https://www.researchgate.net/publication/2406712_The_Insecurity_of_the_Elliptic_Curve_Digital_Signature_Algorithm_with_Partially_Known_Nonces

\subsection{Recover $d$ from a Pair of Signatures with Small $k_1, k_2$}

\textbf{Problem setup}.
Let $p = \texttt{0xffffffffffffd21f}$, and consider an elliptic curve $E: y^2 = x^3 + 3$ over $\F_p$ with the generator
$G = (\texttt{0xcc3b3d1a0c4938ef}, \\ \texttt{0x4ab35ff66f8194fa})$ of order $n = \texttt{0xfffffffefa23f437}$.

We analyze two signatures:

\[
    \begin{array}{c}
        (r_1, s_1) = (\texttt{0x269fa43451c5ff3c}, \texttt{0x1184ec0a74d4be7c}) \\
        \text{and hash } h_1 = \texttt{0xb526aef1a341cfe6}
    \end{array}
\]

\[
    \begin{array}{c}
        (r_2, s_2) = (\texttt{0xf77cda14f5bf50a2}, \texttt{0xcd1143ccc1516b02}) \\ 
        \text{and hash } h_2 = \texttt{0x84768ddee659efea}
    \end{array}
\]

Both signatures use a 32-bit nonce $k$, whereas $n$ is a 64-bit number. We can establish a system of equations:

\[
    \begin{array}{c}
        s_1 \equiv \displaystyle\frac{h_1 + dr_1}{k_1} \mod n \\ \\ 
        s_2 \equiv \displaystyle\frac{h_2 + dr_2}{k_2} \mod n
    \end{array}
\]

Although $d, k_1, k_2$ are unknown, we can express:

\[
    d = \displaystyle\frac{s_1k_1 - h_1}{r_1}
\]

and rewrite the equation as:

\[
    k_1 - s_1^{-1}s_2r_1^{-1}k_2 + s_1^{-1}r_1h_2r_2^{-1} - s_1^{-1}h_1 \equiv 0 \mod n
\]

To simplify, we denote $t = - s_1^{-1}s_2r_1^{-1}k_2$ and $u = s_1^{-1}r_1h_2r_2^{-1} - s_1^{-1}h_1$. This leads to:

\[
    \begin{array}{c}
        k_1 + tk_2 + u \equiv 0 \mod n \\
        \text{or} \\
        -k_1 = tk_2 + u - xn \text{   for some $x$}
    \end{array}
\]

Given that $|k_1|, |k_2| < K = 2^{32}$.

\vspace*{10px}

\textbf{Lattice construction}. We construct the lattice $B = (b_0, b_1, b_2)$:

\[
    B = 
    \begin{pmatrix}
        n & 0 & 0 \\
        t & 1 & 0 \\
        u & 0 & K
    \end{pmatrix}
\]

Note that the vector $v = (-k_1, k_2, K)$ lies within the lattice because:

\[
    v = -xb_0 + k_2b_1 + b_2 = (-xn + tk_2 + u = -k_1, k_2, K)
\]

for some integers $x, k_2$.

\vspace*{10px}

Can we ascertain that $v$ is a short vector discoverable through LLL?

We have:

\begin{center}
    $\lVert v \rVert = \sqrt{k_1^2 + k_2^2 + K^2} \le \sqrt{3K^2} = \sqrt{3}K $
\end{center}

The expected length of the shortest vector is:

\[
    \begin{array}{c}
        \approx 2^{\frac{1}{2}}(\det B)^{\frac{1}{3}} \\
        \approx 2^{\frac{1}{2}}(nK)^{\frac{1}{3}}
    \end{array}
\]

Thus, the criterion:

\begin{center}
    $\lVert v \rVert \le \sqrt{3}K < \sqrt{2}(nK)^{\frac{1}{3}}$
\end{center}

is satisfied when $K < (nK)^{\frac{1}{3}}$ or $K < \sqrt{n}$.

Applying LLL to the second row, we extract:

\begin{center}
    $v = (-k_1, k_2, K) = (\texttt{-0x50a65330, 0x1f5b977a, 0x100000000})$
\end{center}

To recover $d$, we compute:

\[
    d = r_1^{-1}(k_1s_1 - h_1) = \texttt{0xf00e5fb275bfd304}
\]

\subsection{Recover $d$ from Many Signatures with Small $k_i$}

Suppose we have multiple signatures $(r_1, s_1), \ldots, (r_m, s_m)$ along with their corresponding message hashes $h_1, \ldots, h_m$. We can express the relationship:
$s_i \equiv k_i^{-1}(h_i +  dr_i) \mod n$ for $i = 1,\ldots,m$, and eliminate $d$ as previously to obtain:

\[
\begin{array}{c}
    k_1 + t_1k_m + u_1 \equiv 0 \mod n \\
    k_2 + t_2k_m + u_2 \equiv 0 \mod n \\
    \vdots \\
    k_{m-1} + t_{m-1}k_m + u_{m-1} \equiv 0 \mod n
\end{array}
\]

\vspace*{10px}

We then construct the lattice $B$ as follows:

\[
    B = 
    \begin{pmatrix}
     n &  \\
     & n \\
     & & \ddots \\
     & & & n \\
     t_1 & t_2 & \cdots & t_{m-1} & 1 & 0 \\
     u_1 & u_2 & \cdots & u_{m-1} & 0 & K
     \label{eq:lattice_ecdsa}
    \end{pmatrix}
\]

\vspace*{10px}

This lattice includes the vector $v = (-k_1, -k_2, \ldots, k_m, K)$, which can be identified using LLL.

\subsection{Recover $d$ Knowing MSB Bits of Each $k_i$}

What if the most significant bits (MSBs) of each $k_i$ are known?

In such cases, we express $k_i = (a_i + b_i)$, where $a_i$ is the known part, and $b_i$ is the unknown part satisfying $|b_i| < K$. Incorporating these values into the equivalences yields:

\[
    \begin{array}{c}
        k_i + t_ik_m + u_i \equiv 0 \mod n \\
        (a_i + b_i) + t_i(a_m + b_m) + u_i \equiv 0 \mod n \\
        b_i + t_ib_m + a_i + t_ia_m + u_i \equiv 0 \mod n \\
    \end{array}
\]

We then define $u_i' = a_i + t_ia_m + u_i$ as the new term in the lattice \ref{eq:lattice_ecdsa} in place of $u_i$. Consequently, the vector:

\begin{center}
    $v = (-b_1, -b_2, \ldots, b_m, K)$
\end{center}

resides within the lattice and can be detected using LLL. To restore the original nonces $(k_1, \ldots, k_m)$, we simply add the vector $w = (a_1, \ldots, a_m)$ to $v$.

Without specific optimizations, we successfully retrieved the nonces using only the $8$ MSB bits on the curve \texttt{secp256r1} given $100$ signatures, achieving over a $50\%$ success rate with $7$ MSB bits known.

Further refinements to enhance the efficacy of this attack, allowing for fewer known bits, have been demonstrated in the findings by Albrecht and Heninger \cite{ecdsa_result}.

% \subsection{Recover $d$ from many signatures with small $k_i$}
% 
% Suppose we have many signatures $(r_1, s_1), \ldots, (r_m, s_m)$ with message hashes $h_1, \ldots, h_m$. We can write the equivalence
% $s_i \equiv k_i^{-1}(h_i +  dr_i) \mod n$ for $i = 1,\ldots,m$ and we can remove $d$ as before to get
% 
% \[
% \begin{array}{c}
%     k_1 + t_1k_m + u_1 \equiv 0 \mod n \\
%     k_2 + t_2k_m + u_2 \equiv 0 \mod n \\
%     \vdots \\
%     k_{m-1} + t_{m-1}k_m + u_{m-1} \equiv 0 \mod n
% \end{array}
% \]
% 
% \vspace*{10px}
% 
% And create the lattice $B$ as
% 
% \[
%     B = 
%     \begin{pmatrix}
%      n &  \\
%      & n \\
%      & & \ddots \\
%      & & & n \\
%      t_1 & t_2 & \cdots & t_{m-1} & 1 & 0 \\
%      u_1 & u_2 & \cdots & u_{m-1} & 0 & K
%      \label{eq:lattice_ecdsa}
%     \end{pmatrix}
% \]
% 
% \vspace*{10px}
% 
% Same as before this lattice contains $v = (-k_1, -k_2, \ldots, k_m, K)$ that can be found using LLL.
% 
% \subsection{Recover $d$ knowing MSB bits of each $k_i$}
% 
% What if we know the firsts MSB bits of each $k_i$?
% 
% Well, we can write $k_i = (a_i + b_i)$, where $a_i$ is the known part and $b_i$ is the unkown part that satisfies $|b_i| < K$. If we plug these values
% into the equivalence we obtain
% 
% \[
%     \begin{array}{c}
%         k_i + t_ik_m + u_i \equiv 0 \mod n \\
%         (a_i + b_i) + t_i(a_m + b_m) + u_i \equiv 0 \mod n \\
%         b_i + t_ib_m + a_i + t_ia_m + u_i \equiv 0 \mod n \\
%     \end{array}
% \]
% 
% And we can set $u_i' = a_i + t_ia_m + u_i$ to be the value inside the lattice \ref{eq:lattice_ecdsa} instead of $u_i$. In this way the vector
% 
% \begin{center}
%     $v = (-b_1, -b_2, \ldots, b_m, K)$
% \end{center}
% 
% is in the lattice and could be found using LLL. To recover the original $(k_1, \ldots, k_m)$ we must add to $v$ the vector $w = (a_1, \ldots, a_m)$.
% 
% Without any optimization we were able to recover the nonces with the $8$ MSB bits known on the curve \texttt{secp256r1} given $100$ signatures 
% and with a probability bigger than $50\%$ with $7$ MSB bits known.
% 
% It's possible to optimize the attack to recover the nonces with less known bits as shown in the results obtained by
% Albrecht and Heninger \cite{ecdsa_result}.

\chapter*{Conclusion}

The results from implementing these attacks are satisfactory and applicable to real-world scenarios. The success of these attacks often relies on a combination of probability and heuristic approaches, which are not discussed in detail here. As with other branches of mathematics, not all theorems and properties prove immediately useful upon their introduction. However, they may become valuable later, as illustrated by the Hidden Number Problem, which was crucial for the Raccoon Attack.

The exploration of how to exploit side-channel attacks, such as knowing some bits of $p$ in \textbf{RSA}, has proven fruitful, as evidenced by the \textbf{ROCA} vulnerability. The most commonly used cryptographic protocols are generally considered secure; however, improper implementation can easily introduce vulnerabilities. Side-channel attacks, in particular, can significantly undermine the se
% \chapter*{Conclusion}
% 
% The results obtained from the implementation of the attacks are satisfactory and can be used against real world practical cases.
% There is a fair amount of probability and heuristic to make the attacks works that isn't discussed in detail.
% As with other branch of mathematics not all theorems and properties are proved to be useful the moment they were first introduced
% but can become useful years later as the hidden number problem that become crucial for the Raccoon Attack.
% The study on how to exploit side-channel attacks (ex. knowing some bits of $p$ in \textbf{RSA}) has been proven useful
% with \textbf{ROCA}.
% The most used cryptographic protocol used has been proved secure but it's very easy to implement them wrong and side-channel
% vulnerabilities can heavily compromise the security of the protocols.

\bibliographystyle{plain}

\bibliography{lattice_based_attacks}

\end{document}